Linear and Differential Cryptanalysis of DES

### 2009-02-02 trunk/talk/draft1.tex
A couple of small changes to DES Linear slides

\begin{frame}
We perform the attack on the 4-round DES similarly to how we performed the attack on 4-round SPN, by ``undoing'' the last round.\\
\vspace{5mm}
\pause
Note that in the last round, $U_{4R}$ bits are not changed, only $U_{4L}$ bits are affected. Note also that there is only one $U_{4L}$ bit the 3-round equation.
\begin{align*} 
P_L[25] \oplus P_L[14] \oplus P_L[8] \oplus & P_L[3] \oplus U_{4R}[25] \oplus U_{4R}[14] \oplus U_{4R}[8] \\ & \oplus U_{4R}[3] \oplus P_R[17] \oplus U_{4L}[17] = K_1[26] \oplus K_3[26]
\end{align*}
\vspace{5mm}
\pause
This bit, $U_{4L}[17]$, is affected by the result of S-Box 1 in round 4. If we guess correctly for the 6 partial subkey bits $K_{4,1}$ to $K_{4,6}$, the 3-round equation will hold with high bias, otherwise it will likely hold with close to 0 bias.\\
\vspace{5mm}
\pause
Thus, this attack will give us $6$ of the $56$ key bits. \pause What this means in terms of time saved over exhaustive search is this: If it takes a year to find a key through exhaustive search of $2^{56}$ bits, after this attack it will only take $5$ days to exhaustively search the remaining $2^{50}$ bits.
\end{frame}

>>>>

\begin{frame}
\begin{columns}

\begin{column}{0.35\textwidth}
\begin{figure}
\includegraphics[totalheight=0.8\textheight]{DES_4.PNG}
\end{figure}
\end{column}

\begin{column}{0.65\textwidth}
\small
We perform the attack on the 4-round DES similarly to how we performed the attack on 4-round SPN, by ``undoing'' the last round.\\
\vspace{3mm}
\pause
Note that in the last round, $U_{4R}$ bits are not changed, only $U_{4L}$ bits are affected. Note also that there is only one $U_{4L}$ bit the 3-round equation.\\
\begin{align*} 
& P_L[25]    \oplus P_L[14]    \oplus P_L[8]    \oplus P_L[3] \oplus  \\
& U_{4R}[25] \oplus U_{4R}[14] \oplus U_{4R}[8] \oplus U_{4R}[3]  \oplus \\ 
& P_R[17]    \oplus U_{4L}[17] = K_1[26] \oplus K_3[26] \\
\end{align*}\\
\vspace{3mm}
\pause
This bit, $U_{4L}[17]$, is affected by the result of S-Box 1 in round 4. If we guess correctly for the 6 partial subkey bits $K_{4,1}$ to $K_{4,6}$, the 3-round equation will hold with high bias, otherwise it will likely hold with close to 0 bias.\\
\vspace{3mm}
\pause
Thus, this attack will give us $6$ of the $56$ key bits. If it takes a year to find a key through exhaustive search of $2^{56}$ bits, it only takes $5$ days to exhaustively search $2^{50}$ bits.
\end{column}

\end{columns}
\end{frame}
